Variable magnetic field and adaptive mixing-length to reproduce Li abundances and rotation periods of solar-type stars in clusters

### 2024-02-28 project/overleaf/paper2/mblithium.tex
Update the CZ figure eliminating the core size track

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances II]{Variable magnetic field and mixing-length convection parameter to reproduce Li abundances and rotation period of the Sun and solar-type stars}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. Caballero Navarro et al.]{
R. Caballero Navarro,$^{1}$\thanks{E-mail: rcaballeron@correo.ugr.es}
A. Garc\'ia Hern\'andez,$^{1,2}$\thanks{E-mail: agh@ugr.es}
J.~C. Su\'arez$^{1,2}$\thanks{E-mail: jcsuarez@ugr.es}
\\
% List of institutions
% Affiliations should be in the format ‘Department, Institution, Street
% Address, City and Postal Code, Country’.
$^{1}$Dept. Theoretical Physics and Cosmology, University of Granada (UGR), 18071, Granada, Spain\\
$^{2}$Instituto de Astrof\'isica de Andaluc\'ia (CSIC), Glorieta de la Astronom\'ia S/N, 18008, Granada, Spain\\
}

@@

% Abstract of the paper
\begin{abstract}
Finding the reasons why lithium (Li) surface abundance in the Sun and young stellar globular clusters are seemingly anomalous in present-day scenarios, will deliver a better understanding about the mechanisms which influence the Li depletion during the stellar evolution, in particular during the pre-main sequence (PMS) and main sequence (MS).\par

The present work documents an approach in which neither the magnetic field strength nor the MLT parametrisation remains fixed throughout the evolution of the star. The influence of free parameters associated with $\amlt$ and $B$ significantly condition the the Li abundance. We document an alternative to prefixing values to these parameters which ends up providing results that are compatible with observational data for Li abundances of solar siblings hosted in young clusters.\par

We explore, by simulating several grids of PMS and MS rotating models, how the effects of rotational mixing and of the rotational hydrostatic effects governs the Li abundances. The combined effects of the MLT and MB constraint the Li evolution of the solar models. We calculate accurate solar models for which the $\amlt$ and $B$ values are variable over time, i.e. not remaining fixed through the simulations, and make them dependent on important solar parameters. The data obtained from simulations are confronted with observational data measured on young open clusters collected by the Gaia-ESO Survey (GES). The results show how faster rotating models destroy more Li in the late PMS and during the MS, and how the effects of magnetic braking mitigate this destruction. The interconnection between the star's internal convective motions, effective temperature, surface gravity and rotational velocity, as well as its chemical composition, determine the strength of the magnetic field and impact the effectiveness of the convective motions involved in the destruction of Li.

@@

\section{Introduction} \label{sec_intro}
The angular momentum loss (AML) is considered to be governed by magnetically-coupled stellar winds which exert a braking force on the star that eventually causes a loss of angular momentum \citep{UdDoula2002}.

Over the last few years, decades, data from different observational campaigns have allowed studies of the rotational evolution of low and medium-mass stars \citep{Hartman2010,Gallet2013,Bouvier2016}. 
Among them we find studies \citep{Ud-Doula2008,Cranmer2011,Gallet2015,Amard2016} that focus on modelling the effects of magnetic braking (MB) and how its coupling with stellar winds impact the rotational evolution of the star. These models rely on a number of parameters which are fixed, among them the magnetic field strength B (we will come back to this point later). This approach is not sufficient to explain the observational data, and interaction with other processes in the angular velocity evolution of low-mass stars is necessary: the distribution of angular momentum loss in the interior of the star along the pre-main sequence (PMS) and main sequence (MS) \citep{Charbonnel2005, Eggenberger2008, Eggenberger2009, Navarro2020}, and the star - protostellar disc interaction during the PMS \citep{Bouvier2008,Gallet2013,Eggenberger2012,Zanni2012,Bouvier2016}.\par

Discrepancies in Li abundances have been detected in young cluster for which there is as yet no coherent explanation \citep[see][and references therein]{Navarro2020}. Li is destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches $\tli \approx 2.5 x 10^6\; K$. The solar-type stars are known to have a convection zone (CZ) that covers much of the stellar radius during the PMS. In this way, the convective movements that occur in its interior manage to drag the Li to areas in which the BCZ limit exceeds the $\tli$ and Li is burned \citep{Iben1965}. This process slows down as the star approaches the zero-age MS (ZAMS) when the convection zone retreats and moves away from areas where the internal temperature is high enough to destroy the Li.\par

The impact of rotation both on PMS and Li depletion for solar-type stars has already extensively debated in the past \citep{Pinsonneault1997,Jeffries2004,Somers2014} and revised more recently on the basis of the availability of more accurate measures \citep{Gallet2013,Bouvier2016, Bouvier2018, Franciosini2022}. Of particular interest for this paper are the Li observational data measured on young open clusters. The Gaia-ESO Survey (GES) \citep{Gilmore2012,Randich2013,Randich2022} has produced very high quality spectroscopy records of around 110 000 Milky Way stars, in the field and in open clusters, down to magnitude 19, It has delivered good-quality radial velocities and stellar parameters for a large fraction of its unique target stars. Elemental abundances were derived for up to 31 elements for targets observed with the Fibre Large Array Multi Element Spectrograph (FLAMES) and the UV-Visual Echelle Spectrograph (UVES). Lithium abundances are delivered for about 1/3 of the sample.\par

As indicated by its name, GES intends to complement the data about parallaxes and proper motions from the Gaia satellite \citep{Mignard2005} with extremely precise information on radial velocities (RVs), Li, and chemistry in general. The first Gaia data release (GDR1) \citep{Brown2016} dated on 2016 and contained information on the first 14 months of operation. Gaia data release 2 (GDR2) \citep{Brown2018}, released on 2018, contained positions, parallaxes, and proper motions for about 1.3 billion sources, together with photometry. The third data release of Gaia, the Gaia early data release 3 (GEDR3) \citep{Brown2021} and the Gaia data release 3 (GDR3) \citep{Brown2022}, convey updated and more precise measures. GES enhances them and in its latest release (DR5.0) includes all the astrophysical parameters derived by the Gaia-ESO consortium for 114 324 stars \citep{Gilmore2022}. The interest for the present work are:
\begin{itemize}
    \item The radial and projected RVs.
    \item Stellar parameters (effective temperature, surface gravity and metallicity).
    \item Abundances of several elements, among them Li.
    \item Specific parameters for tracing accretion and activity in young stars.
    \item Cluster probability membership.  
\end{itemize}

\section{Data} \label{sec_method}
The data sets provided by the Gaia mission in GEDR3 and GDR3, as well as those published by GES in DR5.0 form the main references used in this work. In general, we focus on the observations made on OCs, and particularly on those analysed in the \cite{Bragaglia2022} and \cite{Randich2022} papers. On the basis of the results presented in these papers, a set of 64 OCs (see Table \ref{tab:oc_full_list}) are selected as the reference dataset for our work.

\begin{table*}
	\centering
	\begin{tabular}{|l l l l l || c c c c c | c c c c c|} 
		\hline
             & & & & & & & $\oomegac$ & & & & & $\oomegac$ & & \\
		Open Cluster & GES\_FLD & Age (Gyr) & [Fe/H] & $N_*$ & 0.095 & 0.10 & 0.105 & 0.11 & 0.115 & 0.12 & 0.125 & 0.13 & 0.14 & 0.1425\\
		\hline
            Berkeley 21 & Br21 & 2.138 & -0.21 & 744 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 1\\
            Berkeley 39 & Br39 & 5.623 & -0.14 & 899 & 1 & 1 & 1 & 1 & 1 & 1 & 2 & 2 & 2 & 2\\
            IC 2602 & IC2602 & 0.036 & -0.06 & 1836 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0\\
            IC 4665 & IC4665 & 0.033 & 0.01 & 567 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0\\
            Messier 67 & M67 & 3.981 & -0.02 & 131 & 6 & 6 & 6 & 6 & 6 & 6 & 6 & 6 & 6 & 6\\
            NGC 2141 & NGC2141 & 1.862 & -0.04 & 853 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2355 & NGC2355 & 1 & -0.13 & 208 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2420 & NGC2420 & 1.698 & -0.15 & 562 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2425 & NGC2425 & 2.399 & -0.13 & 528 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2451 & NGC2451 & 0.035 & -0.08 & 1656 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 2 & 1\\
            NGC 2516 & NGC2516 & 0.24 & -0.04 & 759 & 0 & 0 & 0 & 1 & 1 & 1 & 3 & 4 & 4 & 3\\
            NGC 3532 & NGC3532 & 0.398 & -0.01 & 1145 & 1 & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 0 & 1\\
            NGC 6005 & NGC6005 & 1.259 & 0.22 & 560 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1\\
            NGC 6259 & NGC6259 & 0.269 & 0.18 & 494 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
            NGC 6281 & NGC6281 & 0.513 & -0.04 & 320 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1\\
            NGC 6405 & NGC6405 & 0.035 & -0.02 & 701 & 1 & 1 & 1 & 0 & 1 & 3 & 2 & 0 & 0 & 0\\
            NGC 6633 & NGC6633 & 0.692 & -0.03 & 1662 & 2 & 2 & 2 & 1 & 1 & 0 & 0 & 0 & 1 & 1\\
            NGC 6709 & NGC6709 & 0.191 & -0.02 & 730 & 2 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1\\
            Trumpler 20 & Trumpler20 & 1.862 & 0.13 & 1213 & 3 & 3 & 3 & 2 & 2 & 2 & 2 & 2 & 2 & 2\\
            \hline
	\end{tabular}
 	\caption{List of selected OCs. For each OC, name, GES denomination, estimated age, metallicity and number of components are listed. The number of OC components selected by the different simulations are shown separately. This information is split into two groups, each corresponding to the set of simulations used in the figures shown in the following sections.}
  	\label{tab:oc_reduced_list}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{l l l l l l l} 
		\hline
            ObjectId & $\teff$(K) & $\gsurf$(dex) & FeH(dex) & ALi(dex) & eALi(dex) & Age(Gyr)\\
		\hline
            8510914\_1157003 & 5843 & 4.42 & -0.03 & 1.8 & 0.08 & 3.981\\ 
            8510991\_1146169 & 5797 & 4.4 & -0.01 & 1.65 & 0.08 & 3.981\\ 
            8512177\_1144050 & 5818 & 4.46 & 0 & 1.99 & 0.06 & 3.981\\ 
            8513941\_1200571 & 5792 & 4.41 & -0.02 & 1.49 & 0.13 & 3.981\\ 
            8515559\_1148381 & 5794 & 4.43 & -0.04 & 1.56 & 0.1 & 3.981\\ 
            8520350\_1147480 & 5875 & 4.47 & -0.03 & 1.21 & 0.24 & 3.981\\ 
            \hline
	\end{tabular}
 	\caption{M67 - Filter setup: $\teff$(K) [5787.4892, 5887.6823], $\gsurf$(dex) [4.3979, 4.4977], $\feh$(dex) [-0.05, 0.05], Age(Gyr) [3.977, 3.985]}
	\label{tab:oc_m67}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{l l l l l l l} 
		\hline
            ObjectId & $\teff$(K) & $\gsurf$(dex) & FeH(dex) & ALi(dex) & eALi(dex) & Age(Gyr)\\
		\hline
            7571111-6048156 & 5084 & 4.54 & 0 & 2.29 & 0.08 & 0.24\\ 
            7595031-6044149 & 5108 & 4.52 & -0.05 & 1.92 & 0.08 & 0.24\\ 
            7595280-6032498 & 5093 & 4.49 & -0.03 & 2.29 & 0.08 & 0.24\\ 
            \hline
	\end{tabular}
 	\caption{NGC2516 - Filter setup: $\teff$(K) [5010.0324, 5111.0095], $\gsurf$(dex) [4.4405, 4.5405], $\feh$(dex) [-0.05, 0.05], Age(Gyr) [0.23964, 0.24036]}
	\label{tab:oc_ngc2516}

@@

\subsection{The selection of clusters}
GES DR5.0 provides information on 114 324 stars corresponding to OCs, see \cite{Gilmore2022} for details. Of interest for the present work are those records that meet the condition:
\begin{itemize}
    \item GE\_CL: observed by GES, OC programme field
    \item GE\_SD\_OC: observed by GES, standard field OC
    \item AR\_CL: ESO Archive observation, OC programme field
    \item AR\_SD\_OC: Archive observation, standard field OC
\end{itemize}

This second filtering reduces the population of stars to be considered to a total of 5 895 records which are distributed among the 58 OCs listed in Table \ref{tab:oc_full_list}.\par

\section{Method} \label{sec_method}
The AML is considered to be governed by magnetically-coupled stellar winds which exert a braking force on the star that eventually causes a loss of angular momentum. In addition, throughout the star's evolution, directly observable parameters such as its rotational velocity, effective temperature, surface gravity or abundance of chemical elements undergo variations. These parameters have an influence, either directly or indirectly, on the angular momentum loss caused by the star's magnetic field. On the other hand, we have stellar convection which in most stellar evolution models is prescribed by MLT. That makes use of the mixing-length scale to express which in turn is proportional to the height of the local pressure scale, and to $\amlt$. $\amlt$ is considered a free-parameter and must be determined by comparing the stellar models with some calibrator, typically our Sun, remaining fixed. Therefore, assuming that the magnetic field strength ($B$) and $\amlt$ remain constant throughout the stellar evolution is a necessary simplification in some scenarios but might not correspond with reality.\par

>>>>

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Rotation, magnetic braking \& Li abundances]{Variable magnetic field and adaptive mixing-length to reproduce Li abundances and rotation periods of solar-type stars in clusters}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[R. Caballero Navarro et al.]{
R. Caballero Navarro,$^{1}$\thanks{E-mail: rcaballeron@correo.ugr.es}
A. Garc\'ia Hern\'andez,$^{1}$\thanks{E-mail: agh@ugr.es}
J.~C. Su\'arez$^{1}$\thanks{E-mail: jcsuarez@ugr.es}
\\
% List of institutions
% Affiliations should be in the format ‘Department, Institution, Street
% Address, City and Postal Code, Country’.
$^{1}$Dept. Theoretical Physics and Cosmology, University of Granada (UGR), 18071, Granada, Spain\\
}

@@

% Abstract of the paper
\begin{abstract}
Investigating the apparent anomalies in lithium (Li) surface abundance observed in the Sun and young stellar globular clusters within contemporary astrophysical contexts holds significant promise for advancing our understanding of the mechanisms influencing Li depletion throughout stellar evolution. This study delves into the intricate interplay between rotational mixing and rotational hydrostatic effects in pre-main-sequence (PMS) and main-sequence (MS) stars by employing simulated grids of rotating models. The Li evolution of solar models is scrutinized under the combined influence of Mixing Length Theory (MLT) and magnetic braking (MB), where the magnetic field strength ($B$) and MLT parameterization ($\amlt$) dynamically evolve with key solar parameters. A novel approach, avoiding fixed values for these parameters, is proposed, yielding results consistent with observational data.\par

We generate and test precise solar models that take into account the dynamic nature of $B$ and $\amlt$ during stellar evolution. To validate these models, we compare them with observational data from young open clusters obtained through the Gaia-ESO Survey (GES). Our results are consistent with previous studies that have separately addressed these approaches. Additionally, we integrate both approaches in our models and examine the combined impact they have on the rotational history and surface Li abundances in solar models. The outcomes of our study are in agreement with previous research and are compatible with observational data.\par

The results indicate that there is a strong minimum limit of 1.133 dex for the surface Li abundances in stars similar to the Sun. This aligns with observations of the Sun and provides insight into the complex balance of physical processes that control the amount of Li in stellar atmospheres. Similarly, our calculations yield theoretical values of $\amlt$ that fall within the range of [1.76, 1.78] obtained in previous studies. Based on these promising findings, our models offer a consistent and comprehensive alternative to those that use fixed values and only consider the variables $B$ and $\amlt$ separately. We have successfully obtained results that, although not an exact match to the Sun's actual measurements, are still consistent with its surface Li abundance, angular velocity, and predicted $\amlt$ values.

@@

\section{Introduction} \label{sec_intro}
Throughout the star's evolution, directly observable parameters such as its rotational velocity, effective temperature, surface gravity or abundance of chemical elements undergo variations. These parameters exert an influence, either directly or indirectly, on the angular momentum loss (AML) caused by the star's magnetic field. On the contrary, we encounter stellar convection motions, which, in most stellar evolution models, are described by the mixing-length theory (MLT). MLT makes use of the mixing length scale which is proportional to the height of the local pressure scale, and to the mixing-length parameter ($\amlt$). This parameter, $\amlt$, is considered a fixed free parameter and must be determined by comparing the stellar models with some reference, in most cases our Sun. Therefore, assuming that the magnetic field strength ($B$) and $\amlt$ remain constant throughout the stellar evolution is a necessary simplification in some scenarios but might not correspond with reality.\par

There are different mechanisms that affect the AML in solar-type stars e.g. tidal forces in binary systems or stellar winds. In close binary systems, angular momentum and kinetic energy can be exchanged between the two stars and their orbits through tidal forces \citep{Li2018}. In the latter case, the magnetically-coupled stellar winds efficiently remove angular momentum,  leading to the spin-down of stars with age. This is due to the braking force exerted by stellar winds, which results in the eventual loss of angular momentum \citep{UdDoula2002}.\par

In recent years, the rotational evolution of low- and intermediate-mass stars has been studied thanks to several observational campaigns \citep{Hartman2010, Gallet2013, Bouvier2016}. Some specific investigations, such as those by \citet{Ud-Doula2008, Cranmer2011, Gallet2015, Amard2016}, focus on modeling the effects of magnetic braking (MB) and its coupling with stellar winds on the rotational evolution of stars. These models rely on a number of parameters which are fixed, among them the magnetic field strength $B$. This approach is not sufficient to explain the observational data, and interaction with other processes in the angular velocity evolution of low-mass stars is necessary: e.g., the distribution of AML in the interior of the star along the pre-main sequence (PMS) and main sequence (MS) \citep{Charbonnel2005, Eggenberger2008, Eggenberger2009, Caballero2020}, and the star - protostellar disc interaction during the PMS \citep{Bouvier2008,Gallet2013,Eggenberger2012,Zanni2012,Bouvier2016}.\par

The impact of rotation on both PMS and Li depletion in solar-type stars has been extensively debated \citep{Pinsonneault1997,Jeffries2004,Somers2014}, with recent revisions based on more accurate measurements \citep{Gallet2013, Bouvier2016, Bouvier2018, Franciosini2022}. Li is destroyed in stellar envelopes when the temperature at the base of the convection zone (BCZ) reaches $\tli \approx 2.5 x 10^6\; \mathrm{K}$. Solar-type stars are known to have a convection zone (CZ) that covers much of the stellar radius during the PMS. In this way, the convective movements that occur in its interior manage to drag Li to areas in which the BCZ limit exceeds $\tli$ and Li is burned \citep{Iben1965}. This process slows down as the star approaches the zero-age MS (ZAMS) when the convection zone retreats and moves away from areas where the internal temperature is high enough to destroy the Li. Discrepancies in Li abundances have been detected in young cluster for which there is as yet no coherent explanation \citep[see][and references therein]{Caballero2020}.\par

The Gaia-ESO Survey (GES) \citep{Gilmore2012,Randich2013,Randich2022} has produced very high quality spectroscopy records of around 110 000 Milky Way stars, in the field and in open clusters (OC), down to magnitude 19. It has delivered good-quality radial velocities and stellar parameters for a large fraction of its unique target stars. Elemental abundances for up to 31 elements were derived for targets observed with the Fiber Large Array Multi-Element Spectrograph (FLAMES) and the UV-Visual Echelle Spectrograph (UVES). Li abundances are delivered for about 1/3 of the sample.\par

The primary goal is to explore how MB contributes to the evolution of Li in the Sun and other solar-type stars. We aim to develop evolutionary stellar models in which both $B$ and $\amlt$ are free parameters that vary over time based on specific stellar parameters. We provide semi-empirical approximations for the calculation of AML resulting from braking caused by variable parameters $B$ and $\amlt$. Simulation data are compared with various stellar parameters obtained from Gaia and GES, allowing for the analysis, validation, and refinement of our models against a large candidate set. This has been implemented as an extension to Modules for Experiments in Stellar Astrophysics \citep[MESA; ][]{Paxton2011, Paxton2013,Paxton2015, Paxton2018, Paxton2019}.\par

\section{Data} \label{sec_data}
As indicated by its name, GES intends to complement the data on parallaxes and proper motions from the Gaia satellite \citep{Mignard2005} with extremely precise information on radial velocities (RVs), Li, and chemistry in general. The first Gaia data release (GDR1) \citep{Brown2016} dated 2016 and contained information on the first 14 months of operation. Gaia data release 2 (GDR2) \citep{Brown2018}, released in 2018, contained positions, parallaxes, and proper motions for about 1.3 billion sources, together with photometry. The third data release of Gaia, the Gaia early data release 3 (GEDR3) \citep{Brown2021} and the Gaia data release 3 (GDR3) \citep{Brown2022}, convey updated and more precise measures. GES enhances them and in its latest release (DR5.0) includes all astrophysical parameters derived by the Gaia-ESO consortium \citep{Gilmore2022}. Of interest for the present work are:
\begin{itemize}
    \item The radial and projected RVs
    \item Stellar parameters (effective temperature, surface gravity and metallicity)
    \item Abundances of several elements, among them Li
    \item Specific parameters for tracing accretion and activity in young stars
    \item Cluster probability membership
\end{itemize}

The data sets provided by the Gaia mission in GEDR3 and GDR3, as well as those published by GES in DR5.0 form the main references used in this work. In general, we focus on the observations made on OCs, and particularly on those analysed in the \citet{Bragaglia2022} and \citet{Randich2022} papers. On the basis of the results presented in these papers, a set of 64 OCs (see Table \ref{tab:oc_full_list}) are selected as the reference dataset for our work.

\begin{table*}
	\centering
	\begin{tabular}{|l l l l || c c c c c c c c c c|} 
		\hline
             & & & & & & & & $\omegaini$ & & & & & \\
		Open Cluster & Age (Gyr) & [Fe/H] & $N_*$ & 0.095 & 0.10 & 0.105 & 0.11 & 0.115 & 0.12 & 0.125 & 0.13 & 0.14 & 0.1425\\
		\hline
            Berkeley 21 & 2.138 & -0.21 & 744 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 1\\
            Berkeley 39 & 5.623 & -0.14 & 899 & 1 & 1 & 1 & 1 & 1 & 1 & 2 & 2 & 2 & 2\\
            IC 2602 & 0.036 & -0.06 & 1836 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0\\
            IC 4665 & 0.033 & 0.01 & 567 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 & 0 & 0\\
            Messier 67 & 3.981 & -0.02 & 131 & 6 & 6 & 6 & 6 & 6 & 6 & 6 & 6 & 6 & 6\\
            NGC 2141 & 1.862 & -0.04 & 853 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2355 & 1 & -0.13 & 208 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2420 & 1.698 & -0.15 & 562 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2425 & 2.399 & -0.13 & 528 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
            NGC 2451 & 0.035 & -0.08 & 1656 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 0 & 2 & 1\\
            NGC 2516 & 0.24 & -0.04 & 759 & 0 & 0 & 0 & 1 & 1 & 1 & 3 & 4 & 4 & 3\\
            NGC 3532 & 0.398 & -0.01 & 1145 & 1 & 1 & 0 & 1 & 1 & 1 & 1 & 1 & 0 & 1\\
            NGC 6005 & 1.259 & 0.22 & 560 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1\\
            NGC 6259 & 0.269 & 0.18 & 494 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
            NGC 6281 & 0.513 & -0.04 & 320 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1\\
            NGC 6405 & 0.035 & -0.02 & 701 & 1 & 1 & 1 & 0 & 1 & 3 & 2 & 0 & 0 & 0\\
            NGC 6633 & 0.692 & -0.03 & 1662 & 2 & 2 & 2 & 1 & 1 & 0 & 0 & 0 & 1 & 1\\
            NGC 6709 & 0.191 & -0.02 & 730 & 2 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 1\\
            Trumpler 20 & 1.862 & 0.13 & 1213 & 3 & 3 & 3 & 2 & 2 & 2 & 2 & 2 & 2 & 2\\
            \hline
	\end{tabular}
 	\caption{List of selected OCs. For each OC, name, estimated age, metallicity and number of components are listed. Additionally, the different $\omegaini$ used in the different simulations are shown, where $\omegaini = \oomegac$, $\Omega$ is the star angular velocity at stellar surface, and $\omegac$ is the surface velocity at the equator of a rotating star where the centrifugal force balances the Newtonian gravity. For each OC entry, it is informed per $\omegaini$ how many OC components whose stellar parameters, after comparing them with the equivalent ones calculated by the simulations, are selected. The sorted-out OC members are used as a reference in the pictures shown in the following sections.}
  	\label{tab:oc_reduced_list}
\end{table*}

\subsection{The selection of clusters}
GES DR5.0 provides information on 114 324 stars corresponding to OCs, see \citet{Gilmore2022} for details. We select those records which satisfy any of the conditions:
\begin{itemize}
    \item GE\_CL: observed by GES, OC programme field
    \item GE\_SD\_OC: observed by GES, standard field OC
    \item AR\_CL: ESO Archive observation, OC programme field
    \item AR\_SD\_OC: Archive observation, standard field OC
\end{itemize}

This second filtering reduces the population of stars to be considered to a total of 5 895 records which are distributed among the 64 OCs listed in Table \ref{tab:oc_full_list}.\par

As benchmark data we take a subset of the observations obtained by GES. This subset consists of those stars that belong to OCs, and among them are "hidden" potential twins of our Sun. To identify the latter, an additional screening is performed, this time using the values given by our models during the simulations for metallicity ($\feh$), effective temperature ($\teff$), and surface gravity ($\gsurf$). Finally we are left with those stars for which the GES observations have measured values congruent with our model results, from which we obtain their Li abundances.\par

\subsection{The selection of closest solar twins}
Once we have pre-selected the stars belonging to the OCs, our interest is focused on identifying those that are most similar to our Sun, i.e. solar twins. This brings the question, which stellar parameters should be considered to classify a star as solar twin? Finding them would help us to figure out the origin of our Sun and and gain information about what conditions prevailed in the OCs in which they were born.\par

Several attempts have already been made to find solar siblings and solar twins. The former refers to stars that formed in the same cluster as the Sun and therefore they should have ages and chemical abundances very similar to those of the Sun \citep[see][and references therein]{Adibekyan2018}. However, according to \citet{Strobel1996} solar twins are considered stars possessing fundamental physical parameters such as $\teff$, $\gsurf$, $\feh$, photometric properties, chemical composition, age, luminosity, rotation, and magnetic fields that are similar, if not identical, to those of the Sun. Locating these solar twins presents a formidable challenge due to strict criteria and their dispersion throughout the Milky Way.\par

It is not unusual that, the applied search methods are oriented to find the candidates based on its kinematic properties, and then their temperatures, metallicities and chemical abundances are compared with those of our Sun. In our study, we extend the selection criteria by adding the surface gravity and the age of the star. These two new parameters represent a fine-tuning of the search process for twin candidates. We used the age of the OCs reported in Table \ref{tab:oc_reduced_list} and the data obtained from the set of MESA simulations initiated with different $\omegaini$, where $\omegaini = \oomegac$, $\Omega$ is the angular velocity of the star at the stellar surface, and $\omegac$ is the surface velocity at the equator of a rotating star where the centrifugal force balances Newtonian gravity. The age derived from these simulations provides a crucial parameter for our analysis. This information enables us to determine the remaining simulation parameters, which we subsequently use for comparison with the values associated with stars belonging to the short-listed OCs.\par

\begin{table}
	\centering
	\begin{tabular}{ll} 
		\hline
		Stellar Parameter & Selection Interval\\
		\hline
        OC Age & $\pm0.1\%$ \, for age < 1 Ga \\
        & $\pm0.15\%$ \, for age $\geq$ 1 Ga \\
        $\teff$ & $\pm 50.0 \, \Kelvin$\\
        $\gsurf$ & $\pm 0.05 \, \dex$\\
        $\feh$ & $\pm 0.05 \, \dex$\\
		\hline
	\end{tabular}
    \caption{Selection parameters and intervals applied during the triage process over the OC's components.}
    \label{tab:sel_params}
\end{table}

From approximately 900 identified stars, we select those whose $\feh$, $\teff$, and $\gsurf$ are in the correspondent intervals defined in Table \ref{tab:sel_params} and according to the following classification process: based on the values given by our simulations for the parameters listed in Table \ref{tab:sel_params}, we proceed to select from the different OCs those stars that have values within the established ranges. Specifically, for the estimated age of a given cluster, we establish an interval of $\pm$0.1 for young OCs (age < 1 Ga) and $\pm$0.15 for older counterparts (age >= 1 Ga) \citep[refer to][for comprehensive reference values]{Cantat-Gaudin2020}. This age-dependent interval serves as the basis for selecting simulation time steps that align with the lower and upper age limits of each cluster. Once these two time steps are found, the associated limit values for $\teff$, $\gsurf$ and $\feh$ are gathered. With these value ranges, we proceed then to select in each of the OCs those stars that have values for these same parameters within them. In this way, we finally obtain the candidate stars to be compared with the simulation results.\par

\begin{table*}
	\centering
	\begin{tabular}{l l l l l l l} 
		\hline
            ObjectId & $\teff$(K) & $\gsurf$(dex) & FeH(dex) & ALi(dex) & eALi(dex) & Age(Gyr)\\
		\hline
            8510914-1157003 & 5843 & 4.42 & -0.03 & 1.8 & 0.08 & 3.981\\ 
            8510991-1146169 & 5797 & 4.4 & -0.01 & 1.65 & 0.08 & 3.981\\ 
            8512177-1144050 & 5818 & 4.46 & 0 & 1.99 & 0.06 & 3.981\\ 
            8513941-1200571 & 5792 & 4.41 & -0.02 & 1.49 & 0.13 & 3.981\\ 
            8515559-1148381 & 5794 & 4.43 & -0.04 & 1.56 & 0.1 & 3.981\\ 
            8520350-1147480 & 5875 & 4.47 & -0.03 & 1.21 & 0.24 & 3.981\\ 
            \hline
	\end{tabular}
 	\caption{M67 - Filter setup: GES star identifier, $\teff$(K) [5787.4892, 5887.6823], $\gsurf$(dex) [4.3979, 4.4977], $\feh$(dex) [-0.05, 0.05], Age(Ga) [3.977, 3.985]}
	\label{tab:oc_m67}
\end{table*}

\begin{table*}
	\centering
	\begin{tabular}{l l l l l l l} 
		\hline
            ObjectId & $\teff$(K) & $\gsurf$(dex) & FeH(dex) & ALi(dex) & eALi(dex) & Age(Gyr)\\
		\hline
            7571111-6048156 & 5084 & 4.54 & 0 & 2.29 & 0.08 & 0.24\\ 
            7595031-6044149 & 5108 & 4.52 & -0.05 & 1.92 & 0.08 & 0.24\\ 
            7595280-6032498 & 5093 & 4.49 & -0.03 & 2.29 & 0.08 & 0.24\\ 
            \hline
	\end{tabular}
 	\caption{NGC2516 - Filter setup: GES star identifier, $\teff$(K) [5010.0324, 5111.0095], $\gsurf$(dex) [4.4405, 4.5405], $\feh$(dex) [-0.05, 0.05], Age(Ga) [0.23964, 0.24036]}\label{tab:oc_ngc2516}

@@

The Tables \ref{tab:oc_m67} and \ref{tab:oc_ngc2516} summarize the subset of components of the OC's M67 and NGC2516 selected during the simulations for models initialised with $\omegaini = 0.14$. Each of the tables lists the GES identifier of the star and $\teff$, $\gsurf$, $\feh$, A(Li) and estimated age ranges used in the simulation.

\section{The modelling} \label{modelling}
\subsection{Modelling magnetic intensity} \label{mod_mbi}
Our modeling approach is based on variations of the magnetic field strength throughout a star's evolution, following the formalism introduced by \citet{Gallet2013}. The adopted model employs a dynamo-type prescription, establishing relationships among the star's angular velocity ($\Omega$), effective temperature ($\teff$), magnetic field strength ($B$), and the mass-loss rate ($\mwind$) induced by the stellar wind. We start by enumerating the most relevant aspects and assumptions made in modelling the evolution of magnetic field intensity. \par

A relevant parameter to characterize the influence of a given magnetic field on the stellar wind is $\Omega$. Assuming that the magnetic field is generated by a dynamo, its strength is proportional to some power of $\Omega$,
\begin{ceqn}
\begin{equation}
    f_*B_* \propto \Omega_*^b \label{eq:mf_strenght}
\end{equation}
\end{ceqn}
where $f_*$ is the filling factor, i.e. the fraction of the star surface which is magnetized, $B_*$ is the strength of the magnetic field, $\Omega_*$ is the star angular velocity at stellar surface, and $b$ is the dynamo exponent \citep{Gallet2013}. The multiplication of $f_*$ by $B_*$ will allow us to obtain the mean surface magnetic field. \par
